---
title: "Sum of Transaction to Different Parties in Different States "
output: html_document
Team Members: Zihan Ye(zy2293), Jiachen Xu(jx2318), Jingyi Wang(jw3562), Haikun Du(hd2377)

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,cache = TRUE,warning = FALSE)
```
```{r}
knitr::opts_chunk$set(echo = TRUE)
cm <- read.csv(file="C:\\Users\\jiachen xu\\Desktop\\Columbia Course\\Data Visualization\\Homework\\project\\cm.txt", header = FALSE, sep="|")
cm_header <- read.csv(file="C:\\Users\\jiachen xu\\Desktop\\Columbia Course\\Data Visualization\\Homework\\project\\cm_header_file.csv", header = TRUE, sep=",")
colnames(cm) <- colnames(cm_header)
data <- read.csv(file="C:\\Users\\jiachen xu\\Desktop\\Columbia Course\\Data Visualization\\Homework\\project\\filtered_v2.csv", header = TRUE, sep=",")

data_use <- data[,c("CMTE_ID","NAME","CITY","STATE","TRANSACTION_AMT")]
cm_use <- cm[,c("CMTE_ID","CMTE_CITY","CMTE_ST","CMTE_TP","CMTE_PTY_AFFILIATION")]
all <- merge(data_use,cm_use,by = 'CMTE_ID')

library(ggplot2)
library(tidyr)
library(dplyr)
library(plotly)
```

```{r,fig.width=10,fig.height=8}
library(geofacet)
library(ggthemes)
temp2 <- all[,c('CMTE_PTY_AFFILIATION','STATE','TRANSACTION_AMT')]
temp2 <- temp2[temp2$CMTE_PTY_AFFILIATION == 'DEM' |temp2$CMTE_PTY_AFFILIATION == 'REP',]
temp2 <- temp2 %>% group_by(STATE, CMTE_PTY_AFFILIATION) %>% summarise(sum = sum(TRANSACTION_AMT))

name_vec <- c(as.character(unique(temp2$STATE)), as.character(unique(temp2$CMTE_PTY_AFFILIATION)))

nodes <- data.frame(name = name_vec, id = 0:63)

group.colors <- c(DEM = "#333BFF", REP = "#FF0000")
ggplot(temp2, aes(CMTE_PTY_AFFILIATION, sum, fill = CMTE_PTY_AFFILIATION)) +
    geom_col() +
     coord_flip() +
     facet_geo(~ STATE,grid = "us_state_grid2",scales = "free_x") + ylab("Sum of the Transactions")+ xlab("Parties")+ labs(title = "Sum of Transaction to Parties in Different States", fill = "Committee") + theme(axis.text.x = element_blank(),plot.title = element_text( size= 20),axis.title.x = element_text(size= 15),axis.title.y = element_text(size= 15))+scale_fill_manual(values=group.colors)+ scale_colour_colorblind()
```

For easier comparison among states, we draw the geo facet plot. And considering the intention of this graph is to compare the preference of the party among states, the x scales for each state is different, so the variance on the total transactions would not influence the people's analysis. Also, we tested this graph on CVD simulator website (http://www.color-blindness.com/coblis-color-blindness-simulator/), make sure, even colour vision deficiency person could easily separate the two colours we applied.   
  
Based on this graph, we concluded that CA, NY, MA, IL made more donation to Democratic committee. On the contrary, TX, FL, LA, TN made more contributions to the Republican committee. Also, there have some other states did not have the clear tendency, like CO. The visualization results were quite consistent with our background information. Hence, we concluded that individual transactions from this dataset could represent the state political preference.  