---
title: "Campaign Contribution Pattern Research in 2017-2018 Election Cycle - Occupation"
output: html_document
Team Members: Zihan Ye(zy2293), Jiachen Xu(jx2318), Jingyi Wang(jw3562), Haikun Du(hd2377)
---
This page focuses on researching and visualizing compaign contribution pattern in the perspective of "Occupation".

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
                      message = FALSE)
```

### Occupation Overview: Bubble Chart

We first created a bubble chart in order to have a brief overview of occupations' contribution pattern. The bubble charts plot the top 100 occupations with the largest contribution populations and their average transaction amount. Bubble radius represents total transaction amount. Although “retired” population has the largest overall transaction amount, their average transaction amount is around 800$/person.  “Chairman” and “CEO” might have relatively smaller population, their population average transaction amounts are almost $7,000 and $10,000 per person. Most bubbles cluster around the origin, indicating relatively small population and low average transaction amount.

```{r}
#install.packages("plotly")
library(plotly)
data <- read.csv("datafiles/occupation_200.csv")
data$AVERAGE = data$TRANSACTION_AMT/data$TIMES

p <- plot_ly(data,  x = ~AVERAGE,y = ~CONTRIBUTORS,  type = 'scatter', mode = 'markers', size = ~TRANSACTION_AMT, color = ~OCCUPATION, colors = 'Paired',
             marker = list(opacity = 0.5, sizemode = 'diameter',
                           line = list(width = 2, color = '#FFFFFF')),
             hoverinfo = 'text',
             text = ~paste('Occupation:', OCCUPATION, '<br>Total_Transaction_Amt:', TRANSACTION_AMT, '<br>Mean_Transaction_Amt:', AVERAGE)) %>%
  layout(title = 'TOP 100 Common Occupation Contributions',
         xaxis = list(title = 'MEAN TRANSACTION AMOUNT',
                      showgrid = FALSE),
         yaxis = list(title = 'CONTRIBUTOR POPULATION',
                      showgrid = FALSE),
         showlegend = FALSE)
p
```

### Contributions to Parties: Bar Chart

We then want to further understand occupations contribution patterns to different parties. The following faceted bar chart shows top 5 occupations which contributed the most to a certain party. With fixed x scale, it is obvious that transactions to republican party and democratic party are the most. The top 5 occupations contributed to democratic party is “retired”, “attorney”, “not employed”, “CEO”, “president” in descending order. The top 5 occupations contributed to republican party is “retired”, “CEO”, “homemaker”, “president”, “CEO” in descending order. Apparently, there are some occupations generally contribute a lot to different parties, for instance “retired”, “CEO”, “president”. Whereas, some occupations have inclinations of contributing to a specific party, for instance, “attorney”, “homemaker” and etc.

```{r, fig.width=20, fig.height=10}
library(readr)
library(dplyr)
occuParty <- read.csv("datafiles/occu_to_party.csv")

ggplot(occuParty, aes(x = OCCUPATION, y = TRANSACTION_AMT)) + geom_bar(stat = "identity") + facet_wrap(~CMTE_PTY_AFFILIATION, scales = "free_y") + coord_flip() +
  ggtitle("Top5 Occupations Contribution to Parties") + theme(plot.title = element_text(size = 40, face = "bold"))


```


Additionally, we further investigated into occupations contribution patterns of each state for republican party and democratic party. 
The following graph is a visualization for democratic party. The discovery is consistent of the previous bar plot. In democratic party’s map graph, “retired”, “attorney”, “CEO”, “not employed” appear to the most dominant occupations for most states. There are some other occupations that only donate more in certain states, for instance, “consultant” is in top 5 occupations in DC but not in other staetes. With fixed x, y scales, it is obvious to see that lots of contributions happened in California, New York.


```{r, fig.width=20, fig.height=25}
#install.packages("geofacet")
library(geofacet)
occuDEM <- read.csv("datafiles/occu_to_DEM.csv")
ggplot(occuDEM, aes(OCCUPATION, TRANSACTION_AMT, fill = OCCUPATION)) +
    geom_col() +
     coord_flip() +
     facet_geo(~ STATE,grid = "us_state_grid2") + xlab("Occupation") + ylab("Transaction Amount") + labs(title = "Top 5 Occupations' Contributions to Democratic Party/State") + theme(axis.text.x = element_blank(),plot.title = element_text( size= 35),axis.title.x = element_text(size= 20),axis.title.y = element_text(size= 20))
```



In republican party’s map graph, “retired”, “CEO”, “president” appear to be the most dominant occupations for most states. There are some variations in occupations types in different states.  For instance, “physician” contribute the most in TN. With fixed x, y scales, it is obvious to see that lots of contributions happened in California, Texax and Florida.

```{r, fig.width=20, fig.height=25}
#install.packages("geofacet")
library(geofacet)
occuREP <- read.csv("datafiles/occu_to_REP.csv")
ggplot(occuREP, aes(OCCUPATION, TRANSACTION_AMT, fill = OCCUPATION)) +
    geom_col() +
     coord_flip() +
     facet_geo(~ STATE,grid = "us_state_grid2") + xlab("Occupation") +ylab("Transaction Amount")+ labs(title = "Top 5 Occupations' Contributions to Republican Party /STATE") + theme(axis.text.x = element_blank(),plot.title = element_text( size= 35),axis.title.x = element_text(size= 20),axis.title.y = element_text(size= 20))
```