---
title: "Campaign Contribution Pattern Research in 2017-2018 Election Cycle - Occupation"
output: html_document
Team Members: Zihan Ye(zy2293), Jiachen Xu(jx2318), Jingyi Wang(jw3562), Haikun Du(hd2377)
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
                      message = FALSE)
```

This bar plot reflected many information contribution patterns in occupation perspective and data quality. "RETIRED" and "NOT EMPLOYED" has top 2 largest populations among contributors. This indicates that these two populations overall high concern to the political campaign. Also in data preprocessing, we clustered contributors with occupations, such as "RETIRED-DIRECTOR", "RETIRED-TEACHER" and etc, into "RETIRED". This might contribute to "RETIRED" groups high population.

Also we labeled missing values in occupation attribute as NONE. This group ranked forth largest population. Through our lateral analysis, we found no specific missing patterns in missing values. Thus, there is no way we can find these records' true occupations.

Other occupations with large populations of contributors are "ATTONEY", "PHYSICIAN", "PRESIDENT" and "ENGINEER". These occupations are normally considered as high-salary occupations. It is not surprising that their highly active in compaign contributions.

```{r, fig.width=20, fig.height=10}
library(readr)
library(dplyr)
library(ggplot2)
occu <- read.csv("datafiles/occumissing.csv")

occu$occu <- factor(occu$occu, levels = occu$occu[order(occu$missing)])

g <- ggplot(occu, aes(x = occu , y = missing)) + geom_bar(stat = "identity")  + geom_text(aes(label=missing), hjust = -0.5, vjust=0.5, color = "red") + coord_flip()
g +labs(title="TOP 100 Most Common Contributors' Occupations ",
        x ="Occupations", y = "Frequency/(Count)") + theme(plot.title = element_text(size = 30, hjust = 0.5), axis.title.x = element_text(size= 20),axis.title.y = element_text(size= 20), axis.ticks.x=element_blank(), axis.text.x=element_blank())
```